Stability and Chaos in Multi-Agent Reinforcement Learning

### 2020-10-08 aamas2021-formatting-instructions/aamas21-formatting-instructions/sample.tex
Edit Section 4

%\documentclass[sigconf]{aamas} 
\documentclass[sigconf,anonymous]{aamas}

@@

    \begin{figure}[t]
        \centering

        \begin{subfigure}[b]{0.9 \textwidth}
        \includegraphics[width=0.5\textwidth]{Figures/DynamicalBehavioursTop.png}
        \end{subfigure}

        \begin{subfigure}[b]{0.9 \textwidth}
        \includegraphics[width=0.5\textwidth]{Figures/DynamicalBehaviourschaos.png}
        \end{subfigure}

        \caption{ \label{fig::DynamicalBehaviours} Different types of dynamical behaviour
       displayed
        by learning agents. Here the $x$-axis (resp.~$y$-axis) is the probability with which agent 1 (resp.~agent 2) chooses a given action. a) Figure drawn from \cite{Tuyls2006AnGames}.
        Convergence
        to a unique fixed point in the upper right corner (1, 1). This fixed point is unique, in
        that all trajectories, regardless of initialisation, will converge to this point. b) Limit Cycle, the trajectories converge to cyclic behaviour c,
        d) Chaotic behaviour, here small deviations in the initial conditions can grow
        exponentially. b-d drawn from \cite{Sanders2018}.}
    \end{figure}

@@

\begin{eqnarray}
%    \begin{split}
    0 & = & \xfixed [ \alpha^2 \ttau^2 \Gamma \xfixed q^{p-2} \chi + \nonumber \\
      &  & + \sqrt{2} \alpha \ttau q^{(p-1)/2}z + \sqrt{2} \alpha \htau q^{p/2} z^{p/p-1} + \talpha \rho]
    \label{eqn::fixed_point}
%    \end{split}
\end{eqnarray}
%
where $\chi = \int dt' G(t - t')$, $\eta_0(t) = q^{(p-1)/2}z$, $z$ is
drawn from a Gaussian of zero mean and unit variance. By using (\ref{eqn::fixed_point}) we can
%his gives us an expression for
calculate $\xfixed$. We disregard the choice $\xfixed = 0$, since we
assume that no action will be chosen with exactly zero probability, an
intuitive assumption which is also considered in \cite{Sanders2018}
and \cite{Coolen2005}. The expression inside the squared bracket
admits a positive solution only in the region $\Gamma \in [-1, 0]$,
whilst for positive $\Gamma \leq p-1$, the nature of solutions may
not be guaranteed. Therefore, we restrict our analysis to the region
$[-1, 0]$.

@@

We then examine the
long-term behaviour of the perturbation $\xpert$. To do this, we take
the Fourier transform of (\ref{eqn::Linearised}) and analyse its
behaviour at $\omega = 0$. After some manipulation (Section S4), this gives

@@

\subsection{Discussion}
The fixed point condition (\ref{eqn::Final}) yields a number of testable implications, of which we test the validity through in the subsequent section. These implications are as follows.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width = 0.8 \linewidth]{Figures/pVariance.png}
    \caption{Evaluation of (\ref{eqn::Final}) for varying numbers of players $p$. Here, $\alpha$ is fixed at $0.02$, $\tau = 0.05$, $N = 2$ and $\Gamma = -1$. $p$ is varied from 2 to 12. We see that the value obtained by (\ref{eqn::Final}) tends away from zero monotonically.}
    \label{fig:pvariation}
\end{figure}

@@

% \begin{figure}
%     \centering
%     \includegraphics[width = 1.1 \linewidth]{Figures/Theory.png}
%     \caption{Results of numerical experiments. The heatmaps show the value of the Liapunov Exponent determined by the procedure outlined in Section~\ref{sec:exev} scaled by $1 \times 10^5$ for convenience. Lighter values show a higher exponent whilst darker values indicate lower values. Each experiment is run with $\tau = 0.05$. (Top Left) $p = 2, N = 2$, (Top Right) $p = 2, N = 4$, (Mid-Left) $p = 2, N = 10$, (Mid-Right) $p = 2, N = 20$, (Bottom) $p = 3, N = 5$.}
%     \label{fig:theory}
% \end{figure}
% %\fb{Can we arrange Fig.~\ref{fig:theory} in a 2x2 square?}

@@

To verify experimentally the theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q-Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The aim is to determine the regions in which games learnt using Q-Learning converge to an equilibrium. The results of these experiments are shown in Figure \ref{fig:NumericalExperiments}, with parameters chosen to match those in the analytic assessment (Figure \ref{fig:theory}).
% By examination it is found that, in many cases the system oscillates about some equilibrium point in a low dimensional chaotic manner (i.e. the manner in which the action probabilities change is indistinguishable from random motion) \textbf{get some graphs for this}

% The purpose 
% \st{of this experiment} is to determine a heuristic estimate for the Lyapunov exponent, which measures the degree to which initial conditions which start close to one another converge or diverge as the dynamics evolve \cite{Strogatz2000}. The former indicates a stable system (in the sense of Lyapunov \cite{}) whilst the latter signals instability. The procedure for calculating the Lyapunov exponent is given in (\ref{eq::LiapExpo}).

To generate the numerical simulations in Figure~\ref{fig:NumericalExperiments} we used the
following procedure. Note in the following that for any heatmap in which $\tau$ is swept over, the roles of $\alpha$ and $\tau$ are swapped.
\begin{enumerate}
   \item Fix the parameters $\tau, \gamma$. The former is held at 0.05 and the latter at 0.1. For the heatmaps in which $\tau$ is swept over $\alpha$ is held fixed at 0.05.
   \item Initialise values of $\Gamma, \alpha$ (or $\tau$ appropriately). These will be swept over in the experiment.
\item Generate 15 payoff matrices for both agents by sampling from a multi-variate Gaussian 
(variables are the payoff elements) with mean zero and covariance parameterised by $\Gamma$.
\item For each of these payoff matrices, initialise a set of agents with random initial conditions (i.e., random action probabilities).
\item Allow both sets of agents to learn over a maximum of $1.5 \times 10^4$ iterations.
\item Keep track of the action probabilities over a window of 500 iterations. At the end of each window, determine the percentage difference between the maximum and minimum values of each strategy component
\item If the difference is less than 1 \% consider the game converged. Otherwise continue to the next window.
\item If the game reaches $1.5 \times 10^4$ iterations without satisfying the relative distance criterion, consider it to be non-convergent. Determine the fraction of these 15 games which have converged.

% \item After 500 iterations measure the distance $\Vec{\delta_1} = |A_2 - A_1|$
% \item After 10000 iterations measure the distance $\Vec{\delta_n} = |A_2 - A_1|$
% \item Determine \fb{the Lyapunov exponent} 
% \begin{eqnarray*}
% \label{eq::LiapExpo}
%     \lambda_n & = & \max_{i} 10^{-4} ln(\delta_{n, i}/\delta_{1, i})
% \end{eqnarray*}
% for each action $i$ and average this value over all five realisations of the payoff matrix. 
\end{enumerate}

@@

\begin{figure}[t]
    \includegraphics[width = 1.1 \linewidth, center]{Figures/a5e2var.png}
    \caption{Figures showing the trajectory of action probabilities for 2 players with 5 actions trained on a game of $\Gamma = -0.5$ and $\tau = 0.05$. (Left) $\alpha = 0.01$, (Right) $\alpha = 0.05$.}
    \label{fig:AlphaVariation}
\end{figure}

In Figure \ref{fig:AlphaVariation}, we explicitly plot the trajectories taken by the action probabilities as two agents are trained on a game of $\Gamma = -0.5$. In the left figure, we choose $\alpha = 0.01$ whilst in the right we choose $\alpha = 0.05$. This gives a qualitative assessment of the behaviour at both extremes parameters considered. We note that the typical behaviour of the system is that the system first moves towards a fixed point and then vary in a pseudo-random manner within a vicinity of the fixed point. Critically though, the amount of variation in the case $\alpha = 0.05$ is greater than that of $\alpha = 0.01$, and so it does not settle to a fixed point (i.e. is considered unstable).

\begin{figure}[t]
    \includegraphics[width = 0.75 \linewidth, center]{Figures/pVarianceExpt.png}
    \caption{Figure showing the variation $V(t)$ about a fixed point as $p$ is varied from 2 - 6. $N = 2$, $\alpha = 0.02$, $\tau = 0.05$, $\Gamma = [-1, 0]$. The experiment is conducted for 10 choices of $\Gamma$ and the mean (line) is plotted alongside the standard deviation (shaded region).}
    \label{fig:pVarianceExpt}
\end{figure}

@@

To summarise, we have shown that the analytic result (\ref{eqn::Final}) provides a strong assessment for the effect that $\alpha$, $\tau$, $\Gamma$ and $p$ have on the likelihood of convergence of Q-Learning. We show that, at least for negatively correlated payoffs, the strength of correlation plays no role in determining the stability of the system. Furthermore, though (\ref{eqn::Final}) overestimates the region of instability, it accurately conveys that convergence of Q-Learning is unlikely for games with many players and actions. In fact, our experimental results also verify that stability may only be guaranteed for the limiting case of two-player, two-action games.

Finally, we show that the behaviour of the stability line differs from that of EWA and suggest that this shows that a stability analysis is an important mode of analysis for reinforcement learning equations to ensure that parameters are being chosen appropriately to guarantee the safe convergence of learning.

>>>>

% \documentclass[sigconf]{aamas} 
\documentclass[sigconf,anonymous]{aamas}

@@

    \begin{figure}[t]
    \centering
    \includegraphics[width = 0.8 \linewidth]{Figures/DynamicalBehaviours.png}
    \caption{Evaluation of (\label{fig::DynamicalBehaviours} Different types of dynamical behaviour
       displayed
        by learning agents. Here the $x$-axis (resp.~$y$-axis) is the probability with which agent 1 (resp.~agent 2) chooses a given action. (Top Left) Figure drawn from \cite{Tuyls2006AnGames}.
        Convergence
        to a unique fixed point in the upper right corner (1, 1). This fixed point is unique, in
        that all trajectories, regardless of initialisation, will converge to this point. (Top Right) Limit Cycle, the trajectories converge to cyclic behaviour (Bottom)
        Chaotic behaviour, here small deviations in the initial conditions can grow
        exponentially. (Top Right) and (Bottom) drawn from \cite{Sanders2018}.}
\end{figure}

@@

\small{
\begin{eqnarray}
%    \begin{split}
    0  = \xfixed [ \alpha^2 \ttau^2 \Gamma \xfixed q^{p-2} \chi + \sqrt{2} \alpha \ttau q^{(p-1)/2}z + \sqrt{2} \alpha \htau q^{p/2} z^{p/p-1} + \talpha \rho]
    \label{eqn::fixed_point}
%    \end{split}
\end{eqnarray}
}
%
where $\chi = \int dt' G(t - t')$, $\eta_0(t) = q^{(p-1)/2}z$, $z$ is
drawn from a Gaussian of zero mean and unit variance. By using (\ref{eqn::fixed_point}) we can
%his gives us an expression for
calculate $\xfixed$. We disregard the choice $\xfixed = 0$, since we
assume that no action will be chosen with exactly zero probability, an
intuitive assumption which is also considered in \cite{Sanders2018}
and \cite{Coolen2005}. The expression inside the squared bracket
admits a positive solution only in the region $\Gamma \in [-1, 0]$,
whilst for positive $\Gamma \leq p-1$, the nature of solutions may
not be guaranteed. Therefore, we restrict our analysis to the region
$[-1, 0]$.

@@

We then examine the
long-term behaviour of the perturbation $\xpert$. by taking
the Fourier transform of (\ref{eqn::Linearised}) and analysing its
behaviour at $\omega = 0$. After some manipulation (Section S4), this gives

@@

\begin{figure}[t]
    \centering
    \includegraphics[width = 1.1 \linewidth]{Figures/Theory.png}
    \caption{Evaluation of (\ref{eqn::Final}) over a range of $\alpha$ and $\Gamma$ values. Ligher regions indicate where (\ref{eqn::Final}) is close to zero (i.e. close to the stability boundary) whilst darker regions are further away. We vary $\alpha$ in the range $[0.01, 0.05]$ and $\Gamma$ in the range $[-1, 0]$. $\tau = 0.05$. (Top Left) $p = 2, N = 2$, (Top Right) $p = 2, N = 5$, (Bottom Left) $p = 2, N = 20$, (Bottom Right) $p=3, N = 2$.}
    \label{fig:theory}
\end{figure}

The fixed point condition (\ref{eqn::Final}) yields a number of testable implications, of which we test the validity through in the subsequent section. These implications are as follows.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width = 0.7 \linewidth]{Figures/pVariance.png}
    \caption{Evaluation of (\ref{eqn::Final}) for varying numbers of players $p$. Here, $\alpha$ is fixed at $0.02$, $\tau = 0.05$, $N = 2$ and $\Gamma = -1$. $p$ is varied from 2 to 12. We see that the value obtained by (\ref{eqn::Final}) tends away from zero monotonically.}
    \label{fig:pvariation}
\end{figure}

@@

In this section, we describe and discuss the numerical experiments we use to verify the implications described in the discussion of Section \ref{sec::Theory}. We first describe how these experiments were conducted and then, having presented the results in Figures \ref{fig:NumericalExperiments} and \ref{fig:pVarianceExpt}, we discuss the correlation with the theory.

@@

To verify experimentally the theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q-Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The aim is to determine the regions in which games learnt using Q-Learning converge to an equilibrium. The results of these experiments are shown in Figure \ref{fig:NumericalExperiments}, with parameters chosen to match those in the analytic assessment (Figure \ref{fig:theory}).

To generate the numerical simulations in Figure~\ref{fig:NumericalExperiments} we used the
following procedure. Note in the following that for any heatmap in which $\tau$ is swept over, the roles of $\alpha$ and $\tau$ are swapped.
\begin{enumerate}
   \item Fix the parameters $\tau, \gamma$. The former is held at 0.05 and the latter at 0.1. For the heatmaps in which $\tau$ is swept over $\alpha$ is held fixed at 0.05.
   \item Initialise values of $\Gamma, \alpha$ (or $\tau$ appropriately). These will be swept over in the experiment.
\item Generate 15 payoff matrices by sampling from a multi-variate Gaussian 
(variables are the payoff elements) with mean zero and covariance parameterised by $\Gamma$.
\item For each of these payoff matrices, initialise a set of agents with random initial conditions (i.e., random action probabilities).
\item Allow both sets of agents to learn over a maximum of $1.5 \times 10^4$ iterations.
\item Keep track of the action probabilities over a window of 500 iterations. At the end of each window, determine the percentage difference between the maximum and minimum values of each strategy component
\item If the difference is less than 1 \% consider the game converged. Otherwise continue to the next window.
\item If the game reaches $1.5 \times 10^4$ iterations without satisfying the relative distance criterion, consider it to be non-convergent. Determine the fraction of these 15 games which have converged.

@@

\begin{figure}[t]
    \includegraphics[width = 1 \linewidth, center]{Figures/a5e2var.png}
    \caption{Figures showing the trajectory of action probabilities for 2 players with 5 actions trained on a game of $\Gamma = -0.5$ and $\tau = 0.05$. (Left) $\alpha = 0.01$, (Right) $\alpha = 0.05$.}
    \label{fig:AlphaVariation}
\end{figure}

In Figure \ref{fig:AlphaVariation}, we explicitly plot the trajectories taken by the action probabilities as two agents are trained on a game of $\Gamma = -0.5$. In the left figure, we choose $\alpha = 0.01$ whilst in the right we choose $\alpha = 0.05$. This gives a qualitative assessment of the behaviour at both extremes parameters considered. We note that the typical behaviour of the system is that the system first moves towards a fixed point and then varies in a pseudo-random manner within a vicinity of the fixed point. Critically though, the amount of variation in the case $\alpha = 0.05$ is greater than that of $\alpha = 0.01$.

\begin{figure}[t]
    \includegraphics[width = 0.7 \linewidth, center]{Figures/pVarianceExpt.png}
    \caption{Figure showing the variation $V(t)$ about a fixed point as $p$ is varied from 2 - 6. $N = 2$, $\alpha = 0.02$, $\tau = 0.05$, $\Gamma = [-1, 0]$. The experiment is conducted for 10 choices of $\Gamma$ and the mean (line) is plotted alongside the standard deviation (shaded region).}
    \label{fig:pVarianceExpt}
\end{figure}

@@

To summarise, we have shown that the analytic result
(\ref{eqn::Final}) provides a strong assessment for the effect that
$\alpha$, $\tau$, $\Gamma$ and $p$ have on the likelihood of
convergence of Q-Learning. We showed that, at least for negatively
correlated payoffs, the strength of correlation plays no role in
determining the stability of the system. Furthermore, though
(\ref{eqn::Final}) overestimates the region of instability, it
accurately conveys that convergence of Q-Learning is unlikely for
games with many players and actions. In fact, our experimental results
also verify that stability may only be guaranteed for the limiting
case of two-player, two-action games. Once either of these parameters increase, the region of instability increases rapidly.
%
Finally, we showed that the behaviour of the stability line differs
from that of EWA and suggest that this shows that a stability analysis
is an important mode of analysis for reinforcement learning algorithms
to ensure that parameters are being chosen appropriately to guarantee
the safe convergence of learning.
